---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
```

# tbinenglanddataclean 

[![CRAN_Status_Badge](http://www.r-pkg.org/badges/version/tbinenglanddataclean)](https://cran.r-project.org/package=tbinenglanddataclean)
[![Build Status](https://travis-ci.org/seabbs/tbinenglanddataclean.svg?branch=master)](https://travis-ci.org/seabbs/tbinenglanddataclean)
[![codecov](https://codecov.io/gh/seabbs/tbinenglanddataclean/branch/master/graph/badge.svg)](https://codecov.io/gh/seabbs/tbinenglanddataclean)

tbinenglanddataclean is an R package that contains functions and documentation to reproduce clean and munge available TB data in England.


## Installation

You can install tbinenglanddataclean from github with:

```{r gh-installation, eval = FALSE}
# install.packages("devtools")
devtools::install_github("seabbs/tbinenglanddataclean")
```

## Raw data

This package relies on raw data from several sources, these are;

1. An extract of from the [Enhanced Tuberculosis Surveillance System](https://www.gov.uk/government/publications/tuberculosis-tb-in-england-surveillance-data). Access to this data requires an application to Public Health England.
1. Demographic data from 2000, and from 2001 to 2015 from the [Office of National Statistics](https://www.ons.gov.uk/peoplepopulationandcommunity/populationandmigration/populationestimates/datasets/populationestimatesforukenglandandwalesscotlandandnorthernireland) this data can be downloaded freely.
1. Survey information from the Labour Force Survey, as yearly extracts from 2000-2016 for the [April to June survey](https://discover.ukdataservice.ac.uk/catalogue/?sn=5461). Only registered users can download this data. Registration is possible for those at UK institutions. Other access arrangements can be made at request.

## Cleaning and building the datasets

The included vignette contains the code necessary to build all datasets associated with this package. Each function needs to be pointed at the correct raw data. If the default file names/locations are changed then this will also require updating. Contact [me](https://www.samabbott.co.uk) if you have any problems.
